---
title: "Module 1"
author: "Vyanna Hill"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r}
#Any libraries used
library(ggplot2)
library(dplyr)
```
**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r}
head(inc)
```
```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}
#Hi!!! I might be overthinking this question but lets explore non-visual information

#First, let's see all unique industries of the data set, There are in total twenty five industries in the data set
unique(inc$Industry)

#we can find the Industry with the most employees. We can now pair the max employee found in summary above to its industry of Human resources
inc%>%group_by(Industry)%>%summarise(max_employees=max(Employees))%>% arrange(desc(max_employees))

```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
#We want to find out the the 50 states, how many companies are in each
#using a bar plot, we can plot out the distribution of entries by each state and use the coord_flip() to flip the graph for a "potriat oriented screen"
st_count<-inc%>%group_by(State)%>%summarise(count=n())
ggplot(inc,aes(x=State))+geom_bar()+theme_classic()+ coord_flip()
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
# Lets break this question up as there's multiple parts

#Finding the state with the third most companies
#We can use dplyr to group by state and find its total company count with n()
st_count<-inc%>%group_by(State)%>%summarize(count=n())%>%arrange(desc(count))
head(st_count)

#NY is the third state with the most companies (Knew we were in the top three!)
#Now, Lets Find the average employment by industry
#We only want complete cases, so let create a dataset that only have NY entries with complete cases
NY_ind<-inc%>%filter(State=="NY")
NY_ind<-NY_ind%>%filter(complete.cases(.))


#All industries in NY graph saved below,we can use bloxplot() to show the average employee count and its ranges in employees, we limited the graph to 3,000 as IT has the largest outlier which isn't relevant to our avg employement
ggplot(NY_ind,aes(y=Industry,x=Employees))+geom_boxplot()+theme_classic()+xlim(0,3000)+ggtitle("Average employmee count in NY")
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
#we will assume this question is overall and not NY as it is not specify 

#First, let's only keep complete cases
inc<-inc%>%filter(complete.cases(.))

#Let's find the highest rev/emp by industry with a group_by and summarizing the highest revenue possible
top_rev<-inc%>%group_by(Industry)%>%summarise(avg_revEmp=sum(Revenue)/sum(Employees))%>%arrange(desc(avg_revEmp))

#The top three lucrative industries are Computer Hardware, Energy, and Construction with the highest revenue per employee
head(top_rev)

#We can also visualize this distribution of revenue via bar plot
ggplot(top_rev,aes(x=Industry,y=avg_revEmp))+geom_bar(stat = "identity")+theme_classic()+ coord_flip()+ggtitle("Revenue per employee by Industry")
```
